Nitric oxide feedback to ciliary photoreceptor cells orchestrates UV avoidance

### 2023-05-26 Manuscript/Manuscript.Rmd
MS: edit math model

>>>>

To further analyse the dynamics of responses to UV light and formally describe cPRC phototransduction, synaptic connections and NO retrograde signalling, we developed a mixed cellular-circuit-level mathematical model. We used our Ca2+ imagining data of cPRC, INNOS and INRGW cells collected in wild type, *NOS* knockout, and *NIT-GC2* morphant larvae to formulate assumptions that are the basis of the proposed model.

We model a direct c-opsin1-dependent response to UV leading to the initial rise in Ca2+ levels. The mechanism of this signal is not known but may include Galpha and Gbeta signalling, CNG channel activation or other pathways. We further model a decrease following UV stimulation in Ca2+ levels that is dependent on NIT-GC2. The

in cPRC cells increase with cGMP. cGMP decreases under NIT-GC2
exposure to UV (mechanism unknown, comparison of Fig. 4A, B and M with Fig. 4N, described
in section ’NIT-GC1 is required for NO-mediated feedback to cPRCs during the UV response’).
cGMP increases with increasing NO (comparison of Fig. 4A with Fig. 4B described in sections
’NO feedback tunes cPRC responses to UV/violet stimulation’ and ’NIT-GC1 produces cGMP in a
NO-dependent manner’ and ’NIT-GC1 is required for NO-mediated feedback to cPRCs during the
UV response’).
Based on the connectome structure (Fig. 1F) we infer a feedforward coupling between cPRC
cells and IN-NOS and IN-RGW cells. The proposed synaptic coupling is UV-dependent and decays
linearly. It is suppressed in NIT-GC2 morphants (observations in Fig. 5
